Metal Mixing and Ejection in Dwarf Galaxies is Dependent on Nucleosynthetic Source

### 2018-07-29 main.tex

\begin{figure}
\centering
\includegraphics[width=0.95\linewidth]{species_bar_sources.png}\\
\caption{The fraction of total mass in each metal species produced by each of the four possible nucleosynthetic channels in our model. These channels differ in both when metals are ejected, as determined by stellar evolution, and the energy with which they are ejected into the ISM. We note that the minimal contribution from Type Ia supernovae for the iron-peak elements is because only 16 of them have exploded by the end of our 500~Myr simulation.}
\label{fig:species_sources}
\end{figure}

The only physics that separates the dynamical evolution of these elements in our simulations is the individual sources of enrichment: AGB winds (stars less than 8~M$_{\odot}$), stellar winds (stars above 8~M$_{\odot}$), core collapse supernovae, and Type Ia supernovae. These channels differ in: 1) how long after a given star formation event they occur, 2), by consequence, the typical ISM properties in which they occur, 3) how much energy is associated with each event, which determines ejecta temperature and velocity. To understand where each of the elements in Figure~\ref{fig:species_fractions} originated, we show the mass fraction of metals produced through each channel in Figure~\ref{fig:species_sources}. Core collapse supernovae are responsible for over 90\% of the total metal enrichment in our galaxy, but this is clearly not true for all elements. In particular, a majority of N (74\%) and Ba (65\%) are ejected by AGB winds; both are retained at a higher rate than the rest of the metals. That this behavior exists for N and Ba in our simulations is dependent upon the metallicity of our galaxy and choice of stellar yields tables, but we can generalize this result to say that, for any assumed set of yields, low mass galaxies should more easily retain \textit{any} elements synthesized predominately in AGB winds, as compared to elements synthesized through supernovae.\footnote{One might expect that much of the N and Ba that is ejected by galactic winds is comprised mostly of the $\sim$30\% of each species produced through the stellar winds of more massive stars and supernovae. However, this cannot be determined as we lack Lagrangian information about individual gas elements.}
% AE: Actually if there are significant isotope variations between nucleosynthetic sources (i.e. N produced in AGB winds is a different isotope than the trace N produced in SNe) then that could be a very cool thing to distinguish if we had good gas-phase abundance measures of various isotopes... but maybe not if they aren't very stable....
 We discuss how this result may extend towards other metal yields that are not well sampled on our relatively short (500 Myr) simulation timescales in Section~\ref{sec:discussion:metal yields}.

AGB winds have low energy and velocities (10~km~s$^{-1}$) as compared to the energy and typical expansion velocities of supernovae ($\sim$1000~km~s$^{-1}$). In addition, their longer timescales relative to massive stars means that AGB stars are typically removed from their birth regions. The changes in typical ISM density and height of these events contributes to these variations. We show histograms of the average number density within 20~pc of any given enrichment source (top) and height above/below the disk (bottom) within 1~Myr of each event (as limited by our output cadence) in Figure~\ref{fig:spatial distribution}. As shown, SN peak at very low densities, indicating that most explode in superbubbles, regions carved out by previous SNe. AGB stars predominately release their metals close to the average ISM density. The scale height distributions for both events show no significant differences. We do not expect these differences to be the dominant effect in driving the differential evolution of elements ejected by AGB winds vs. supernovae, compared to the energetics, but can certainly play a significant role in determining the mixing behavior of individual enrichment events. The changes to mixing behavior as a function of ISM properties will be investigated in more detail in a future work.

>>>>

\begin{figure}
\centering
\includegraphics[width=0.95\linewidth]{species_bar_sources.png}\\
\caption{The fraction of total mass in each metal species produced by each of the four possible nucleosynthetic channels in our model. These channels differ in both when metals are ejected, as determined by stellar evolution, and the phase of the ISM into which they are ejected.
We note that the minimal contribution from Type Ia supernovae for the iron-peak elements is because  no older stellar population was initialized, so only 16 of them have exploded by the end of our 500~Myr simulation.}

\label{fig:species_sources}
\end{figure}

The only physics that separates the dynamical evolution of these elements in our simulations is the individual sources of enrichment: AGB winds (stars less than 8~M$_{\odot}$), stellar winds (stars above 8~M$_{\odot}$), core collapse supernovae, and Type Ia supernovae. These channels differ in: 1) how long after a given star formation event they occur, 2), by consequence, the typical ISM properties in which they occur, and 3) how much energy is associated with each event, which determines ejecta temperature and velocity. To understand where each of the elements in Figure~\ref{fig:species_fractions} originated, we show the mass fraction of metals produced through each channel in Figure~\ref{fig:species_sources}. Core collapse supernovae are responsible for over 90\% of the total metal enrichment in our galaxy, but this is clearly not true for all elements. In particular, a majority of N (74\%) and Ba (65\%) are ejected by AGB winds; both are retained at a higher rate than the rest of the metals. That this behavior exists for N and Ba in our simulations is dependent upon the metallicity of our galaxy and choice of stellar yield tables, but we can generalize this result to say that, for any assumed set of yields, low mass galaxies should more easily retain \textit{any} elements synthesized predominately in AGB winds, as compared to elements synthesized through supernovae.\footnote{One might expect that much of the N and Ba that is ejected by galactic winds is comprised mostly of the $\sim$30\% of each species produced through the stellar winds of more massive stars and supernovae. However, this cannot be
%mm determined
    proven
as we lack Lagrangian information about individual gas elements.}
% AE: Actually if there are significant isotope variations between nucleosynthetic sources (i.e. N produced in AGB winds is a different isotope than the trace N produced in SNe) then that could be a very cool thing to distinguish if we had good gas-phase abundance measures of various isotopes... but maybe not if they aren't very stable....
 We discuss how this result may extend towards other metal yields that are not well sampled on our relatively short (500 Myr) simulation timescales in Section~\ref{sec:discussion:metal yields}.

AGB winds have low energy and velocities (10~km~s$^{-1}$) as compared to the energy and typical expansion velocities of supernovae ($\sim$1000~km~s$^{-1}$). In addition, their longer timescales relative to massive stars means that AGB stars are typically removed from their birth regions and randomly distributed through the galactic disk. The changes in typical ISM density and height of these events contributes to these variations. We show histograms of the average number density within 20~pc of any given enrichment source (top) and height above/below the disk (bottom) within 1~Myr of each event (as limited by our output cadence) in Figure~\ref{fig:spatial distribution}. As shown, SN peak at very low densities, indicating that most explode in superbubbles, regions carved out by previous SNe. AGB stars predominately release their metals close to the average ISM density. The scale height distributions for both events show no significant differences. We do not expect these differences to be the dominant effect in driving the differential evolution of elements ejected by AGB winds vs. supernovae, compared to the energetics, but can certainly play a significant role in determining the mixing behavior of individual enrichment events. The changes to mixing behavior as a function of ISM properties will be investigated in more detail in a future work.
